---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  warning = FALSE,
  message = FALSE,
  tidy = FALSE,
  fig.align='center',
  comment = "#>",
  fig.path = "man/figures/README-",
  R.options = list(width = 200)
)
```
# netlit: Augment a literature review with network analysis statistics
------

`netlit` provides functions to generate network statistics from a literature review. Specifically, it process a dataset where each row is a proposed relationship ("edges") between two concepts or variables ("nodes").

To install `netlit` from CRAN, run the following:

```{r, eval=F}
install.packages("netlit")
```

## Basic Usage

The `review()` function takes in a data frame, `data`, that includes `from` and `to` columns (a directed graph structure). 

In the example below, we use example data from [this project on redistricting](https://github.com/judgelord/redistricting). See `vignette("netlit")` for more details on this example.

```{r}
library(netlit)

data("literature")

names(literature)
```

The four functions `netlit` offers are `make_edgelist()`, `make_nodelist()`, `augment_nodelist()`, and `review()`. `review()` is the primary function (and probably the only one you need), and the others are helper functions that perform the individual steps that `reviews()` all at once. `review()` takes in a data set with a `from` column, a `to` column, and possibly edge attributes and returns a data structure that includes the inputs augmented with network statistics and an `igraph` graph that can be plotted using functionality in `igraph` or other network visualization packages. We demonstrate `review()` below:

```{r}
lit <- review(literature)

lit
```

Edge and node attributes can be added as well using the `edge_attributes` and `node_attributes` arguments. `edge_attributes` identifies the column in the data set that correspond to edge attributes, and `node-attributes` is a separate dataset contain attributes for each node in the primary data set.

```{r}
data("node_attributes")

head(node_attributes)

lit <- review(literature,
              edge_attributes = c("mechanism", "cites"),
              node_attributes = node_attributes)

lit
```

